Characterizing subarctic biomes for land surface modeling of pollution and climate risk

### 2021-04-18 VegOzone.tex
Spell-checked

\begin{abstract}
Here we assess the importance of characterization of key meteorological variables and subarctic biomes on the modeled susceptibility of subarctic vegetation to ozone used in pollution and climate risk assessments. Future climate changes will manifest in both direct and indirect effects influencing the species' exposure to air pollution. A likely increase in the frequency of persistent heatwaves in the future will inflict heat-related stress while generating favorable conditions for the formation of ground level ozone, e.g. due to wildfires.\\
We have monitored ground level ozone concentrations at a high latitude site in Northern Norway during the growing season 2018/19. Due to data acquisition problems, ozone concentrations for two weeks in July 2018 were missing from our record. We present a reconstruction based on a probability density function (PDF) and concidering the Swedish and Finnish atmospheric monitoring sites. We review available ground level ozone reanalysis products for the region and find only CAMS Regional Air Quality reanalysis ensemble satisfactory.
In terms of anomalies from a multi-annual mean, ozone concentrations were only significantly enhanced in April/May 2018 before the start of the growing season. Temperatures and global irradiance showed consecutively significant positive anomalies during Spring and Summer 2018. In 2019 only early spring and late summer were significantly warmer, while colder during summer. Precipitation was lower in both years but displayed a large variability.\\
We investigate the difference in uptake of ozone using the $\mathrm{DO_3SE}$ model, for typical vegetation at the location. We find that bespoke parameterization for subarctic vegetation is essential for an accurate estimation of Phyto-toxic Ozone Dose (POD) and an assessment of damage risk. Accurate estimates of the start and length of growing season are also crucial (up to $15\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ deviation in $\mathrm{POD_1}$).\\
We find that temperature and its derivative vapor pressure deficit are likely the most important climate variables while the variability of ozone concentration within climatological reasonable limits plays a minor role.
The use of subarctic parameterization suggests an improvement in risk assessment as POD values are increased compared to standard expectations; corroborated by the evidence of visible injury in the Svanhovd ozone garden. We estimate a total biomass reduction in Norway spruce of $2.5-3\,\unit{\%}$, downy birch by $15.5-17.4\,\unit{\%}$ and an above-ground biomass reduction in perennial grassland of $12.4-13.7\,\unit{\%}$.  
\end{abstract}

@@

A substantial body of evidence exists that suggests flux-based metrics, that relate stomatal ozone uptake to vegetation damage, are biologically more relevant for risk assessments than exposure-based metrics. This is because they are able to account particular species characteristics (i.e. physiology and phenology) as well as environmental conditions that can decouple the relationship between ozone concentration, ozone uptake, and consequent damage \citep{PT:Emberson2020}. These flux based metrics have been shown to be better suited to represent the actual risk to vegetation from ozone, especially in the climatically extreme parts of Europe \citep{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017}. Most of these previous studie have focussed on the Mediterranean where soil moisture deficits are thought to be most influential in decoupling concentration from uptake. Relatively few studies have explored the situation in Northern European climates where low temperatures may be considered more likely to influence ozone uptake \citep[e.g.]{iF:Juran2018}.

To estimate the flux-based metric, also referred to as the Phytotoxic Ozone Dose over a threshold y ($\mathrm{POD_y}$), an estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf and environmental factors affecting stomatal conductance as described further in Section~\ref{sec:do3se}. $\mathrm{POD_y}$ is then calculated according to:
%
\begin{equation}
  \mathrm{POD_y} = \int{(\Phi_\mathrm{sto}-y)\cdot \mathrm{dt}},
  \label{eq:pod}
\end{equation}
%
with the hourly averaged stomatal ozone flux $\Phi_\mathrm{sto}$ (see Eq.~(\ref{eq:flux_stomata}) and a stomatal ozone flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}\,s^{-1}}. The flux threshold $\mathrm{y}$ represents the detoxification potential of the plant and is typically only exceeded during daylight hours (i.e. when global radiation is above $50\,\unit{W\,m^{-2}}$). $\mathrm{POD_y}$ is given in \unit{mmol\,m^{-2}} over the growing season.

@@

An application of the Mapping Manual method was made by \citet{GCB:Mills2011} where exceedance of the CL was related to observations of ozone damage to a clover biomonitor. This study clearly demonstrated the improved ability of flux-based metrics (in comparison to exposure-based metrics) to identify the geographical distribution of the risk of ozone damage and also showed that the ozone damage could extend into more northerly regions of Europe. A similar biomonitoring study was conducted at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd to assess whether ozone was capable of inducing damage to vegetation growing under northern Fennoscandia conditions. In 2018, we indeed observed ozone damage on clover and tobacco. In contrast, no such damage was found on the clovers in 2019.

@@

The methods described above rely on accurate representation of plant physiological response to climate variables (i.e. the species or plant functional type-specific responses used to parameterize risk assessment models) and are most often parameterized using data for less extreme climates, i.e., temperate or continental climates; and to a lesser extent, Boreal and Mediterranean climates. Parameterization of these models for more extreme bio-geographical regions such as subarctic climates has not been performed and relies on assuming that less extreme Boreal parameterization will be applicable. In this study, conditions at Svanhovd for the growing seasons 2018/19 shall serve as a reference for present and probable future conditions in northern Fennoscandia.
It is important to understand the implications of using these more generic parameterizations for more extreme climate regions. Here we explore how risk estimates might be over or underestimated using generic vs bespoke subarctic parameterization. We also explore the role of \chem{[O_3]} variability and changing climate variables to understand how this risk might alter in the future with the advent of climate change. This allows an understanding of how the threat from ozone may change over time and which are the key aspects of plant physiology that might determine risk. Hence, we determin which parameters require special attention to define accurately for pollution impact assessment and more generally for describing gas exchange response to changing climates that will influence biochemical cycling. This might be one key to solve long-standing issues of coupled earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface models \citep{GMD:Poulter2015,JAMES:Lawrence2019}.\\

@@

There are two global reanalysis products available at the European Centre for Medium-Range Weather Forecasts (ECMWF) that include atmospheric tracers, in particular ozone, the Monitoring Atmospheric Composition and Climate (MACC) and the newer Copernicus Atmosphere Monitoring Service reanalysis (CAMSRA) \citep{ACP:Inness2013, ACP:Inness2019}. The temporal as well as spatial resolution of these reanalysis products is rather coarse: 3-hourly and $0.75\,\unit{^\circ}\times0.75\,\unit{^\circ}$ or roughly $29.3\,\unit{km}\times 83.4\unit{km}$ at the location of Svanvik. From Copernicus member institutions, higher resolution suface ozone ranalysis data are available for Europe (CAMS Regional Air Quality (CAMSRAQ) reanalysis ensemble). The CAMSRAQ is based on nine European state-of-the-art numerical air quality models. The ensemble mean is at fairly high spatial and temporal resolution compared to the global reanalysis: $0.1\,\unit{^\circ}\times 0.1\,\unit{^\circ}$ or roughly $3.9\,\unit{km}\times 11.1\,\unit{km}$ and 1-hourly, respectively. The periodes covered differ and no data is available before the turn of the millennium. The CAMSRA and CAMSRAQ are available in near real time, but only the former covers a period of sufficient length for climate analysis (2003 -- present). All reanalyses use different but, at that time, the latest versions of the operational weather forecast system (OpenIFS) from ECMWF. They differ substantially in the assimilated observational ozone data. The MACC reanalysis assimilates only satellite-derived tropospheric column ozone, while CAMSRA also includes ozone profiles from satellite retrievals. Most importantly for our study, in situ observations from ozone near-surface station networks are only assimilated in the CAMSRAQ reanalysis ensemble. The quality concerning observations in northern Fennoscandia will be discussed in Section~\ref{subsubsec:clim_ozone}. Due to the found substantial discrepancies between reanalysis data and site observations in northern Fennoscandia, we refrain from using any of the global reanalysis data.

\begin{table}[t]
  \caption{Global/refional ozone reanalysis products used in this study.}
  \label{tab:ozone_rea}
\begin{tabular}{llcclcc}
\tophline
Name & Provider & \multicolumn{2}{c}{Resolution} & Time period & Meteorological forcing & \chem{O_3} assimilation\\
&              & spatial & temporal & & \\
\middlehline
MACC & ECMWF & $0.75^\circ \times 0.75^\circ$ & 3-hourly & 2003 -- 2012 & OPS & satellite $^\triangledown$\\
CAMSRA & ECMWF & $0.75^\circ \times 0.75^\circ$ & 3-hourly & 2003 -- 2012 $^\dagger$ & ERA5 / OPS $^\ddagger$ & satellite $^\blacktriangledown$\\
CAMSRAQ & Copernicus & $0.1^\circ \times 0.1^\circ$ & 1-hourly & 2014 -- 2018 $^\dagger$ & OPS $^\star$ & in situ $^\vartriangle$\\
\bottomhline
\end{tabular}
\belowtable{$^\dagger$ Subset of the data used in this study. The full data set extends into near real-time; $^\ddagger$ ERA5 (2003--2016), OPS (later); $^\star$ EURAD uses WRF for downscaling of operational IFS; $^\triangledown$ MLS, OMI - tropospheric column; $^\blacktriangledown$ SCIAMACHY, MIPAS, MLS, OMI, GOME2, SBUV2 - tropospheric column + profile; $^\vartriangle$ METEO France NRT.} % Table Footnotes
\end{table}

@@

We hypothesize that perennial vegetation will likely prioritize their growth or fast adolescenceand and hence their carbon sequestration, during a short, subarctic growing season. Because the mapping manual version of the $\mathrm{DO_3SE}$ model currently does not simulate $A_\mathrm{n}$, we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$. We further assume that $f_\mathrm{VPD}$ and $f_\mathrm{SWP}$ suffice our biomes. The procedure is graphically illustrated in Fig.~\ref{fig:f_temp_grassland} for grassland. We aim to increase the enclosed area between probability density function (PDF) and the respective response function. We show two climatological reference periods 1990s and 2000s to indicate that the vegetation is subject to ongoing climate change and hence continuous acclimation. Resulting response functions for Norway spruce (Picea abies) and downy birch are displayed in Appendix~Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}.

@@

As pointed out by \citet{ACP:Bueker2012}, a very important factor in deriving a proper $\mathrm{POD_y}$ is the soil texture or how much water the soil can hold. As Svanhovd is located in the bed of the Pasvik river, we assume a sandy loam texture at $60\,\unit{cm}$ depth. The estimate of leaf boundary layer resistence is a function of windspeed which will be influenced by the tree height and leaf dimensions. Leaves of downy birch in northern Fennoscandia are smaller than southern specimens' with leaf dimensions of $(3.0\pm 0.5)\,\unit{cm}$. In addition, \citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area. Both these factors will affect leaf boundary layer resistance, the smaller leaves would tend to reduce leaf boundary layer resistances whilst the shorter trees would reduce canopy roughness and increase canopy boundary layer resistances.

@@

From the difference between 2018 and 2019, we infer a substantial interannual variability. The interannual variability is estimated to be larger than our reported systematic deviations and can be of the same order of magnitude as the effect of the temperature acclimation. This means that the choice of temperature acclimation has the potential to either revoke or amplify the risk of ozone damage for vegetation. However, the risk of ozone damage in downy birch and perennial grassland in comparison with the generic default MM parametrizations is increased, regardlessly.

@@

\begin{acknowledgements}
  Bj{\o}rg Rognerud for processing SeNorge.no data with respect for begin of growing season.
  Tore Flatlandsmo Berglen for hourly pressure data from Svanhovd.
  This work was financed by the Research Council of Norway (NRC) grand No. xxxxx.
\end{acknowledgements}

>>>>

\begin{abstract}
Here we assess the importance of characterization of key meteorological variables and subarctic biomes on the modeled susceptibility of subarctic vegetation to ozone used in pollution and climate risk assessments. Future climate changes will manifest in both direct and indirect effects influencing the species' exposure to air pollution. A likely increase in the frequency of persistent heatwaves in the future will inflict heat-related stress while generating favorable conditions for the formation of ground level ozone, e.g. due to wildfires.\\
We have monitored ground level ozone concentrations at a high latitude site in Northern Norway during the growing season 2018/19. Due to data acquisition problems, ozone concentrations for two weeks in July 2018 were missing from our record. We present a reconstruction based on a probability density function (PDF) and considering the Swedish and Finnish atmospheric monitoring sites. We review available ground level ozone reanalysis products for the region and find only CAMS Regional Air Quality reanalysis ensemble satisfactory.
In terms of anomalies from a multi-annual mean, ozone concentrations were only significantly enhanced in April/May 2018 before the start of the growing season. Temperatures and global irradiance showed consecutively significant positive anomalies during Spring and Summer 2018. In 2019 only early spring and late summer were significantly warmer, while colder during summer. Precipitation was lower in both years but displayed a large variability.\\
We investigate the difference in uptake of ozone using the $\mathrm{DO_3SE}$ model, for typical vegetation at the location. We find that bespoke parameterization for subarctic vegetation is essential for an accurate estimation of Phyto-toxic Ozone Dose (POD) and an assessment of damage risk. Accurate estimates of the start and length of growing season are also crucial (up to $15\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ deviation in $\mathrm{POD_1}$).\\
We find that temperature and its derivative vapor pressure deficit are likely the most important climate variables while the variability of ozone concentration within climatological reasonable limits plays a minor role.
The use of subarctic parameterization suggests an improvement in risk assessment as POD values are increased compared to standard expectations; corroborated by the evidence of visible injury in the Svanhovd ozone garden. We estimate a total biomass reduction in Norway spruce of $2.5-3\,\unit{\%}$, downy birch by $15.5-17.4\,\unit{\%}$ and an above-ground biomass reduction in perennial grassland of $12.4-13.7\,\unit{\%}$.  
\end{abstract}

@@

A substantial body of evidence exists that suggests flux-based metrics, that relate stomatal ozone uptake to vegetation damage, are biologically more relevant for risk assessments than exposure-based metrics. This is because they are able to account particular species characteristics (i.e. physiology and phenology) as well as environmental conditions that can decouple the relationship between ozone concentration, ozone uptake, and consequent damage \citep{PT:Emberson2020}. These flux based metrics have been shown to be better suited to represent the actual risk to vegetation from ozone, especially in the climatically extreme parts of Europe \citep{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017}. Most of these previous studies have focused on the Mediterranean where soil moisture deficits are thought to be most influential in decoupling concentration from uptake. Relatively few studies have explored the situation in Northern European climates where low temperatures may be considered more likely to influence ozone uptake \citep[e.g.]{iF:Juran2018}.

To estimate the flux-based metric, also referred to as the Phyto-toxic Ozone Dose over a threshold y ($\mathrm{POD_y}$), an estimate of the stomatal \chem{O_3} flux ($\Phi_\mathrm{sto}$) is calculated based on the assumption that the concentration of \chem{O_3} at the top of the canopy represents a reasonable estimate of the concentration at the upper surface of the laminar layer for a sunlit upper canopy leaf and environmental factors affecting stomatal conductance as described further in Section~\ref{sec:do3se}. $\mathrm{POD_y}$ is then calculated according to:
%
\begin{equation}
  \mathrm{POD_y} = \int{(\Phi_\mathrm{sto}-y)\cdot \mathrm{dt}},
  \label{eq:pod}
\end{equation}
%
with the hourly averaged stomatal ozone flux $\Phi_\mathrm{sto}$ (see Eq.~(\ref{eq:flux_stomata}) and a stomatal ozone flux threshold $\mathrm{y}$ both given in units of \unit{nmol\,\chem{O_3}\,m^{-2}\,\mathrm{PLA}\,s^{-1}}. The flux threshold $\mathrm{y}$ represents the detoxification potential of the plant and is typically only exceeded during daylight hours (i.e. when global radiation is above $50\,\unit{W\,m^{-2}}$). $\mathrm{POD_y}$ is given in \unit{mmol\,m^{-2}} over the growing season.

@@

An application of the Mapping Manual method was made by \citet{GCB:Mills2011} where exceedance of the CL was related to observations of ozone damage to a clover bio-monitor. This study clearly demonstrated the improved ability of flux-based metrics (in comparison to exposure-based metrics) to identify the geographical distribution of the risk of ozone damage and also showed that the ozone damage could extend into more northerly regions of Europe. A similar bio-monitoring study was conducted at the Norwegian Institute of Bioeconomy Research (NIBIO) Environment Centre Svanhovd to assess whether ozone was capable of inducing damage to vegetation growing under northern Fennoscandia conditions. In 2018, we indeed observed ozone damage on clover and tobacco. In contrast, no such damage was found on the clovers in 2019.

@@

The methods described above rely on accurate representation of plant physiological response to climate variables (i.e. the species or plant functional type-specific responses used to parameterize risk assessment models) and are most often parameterized using data for less extreme climates, i.e., temperate or continental climates; and to a lesser extent, Boreal and Mediterranean climates. Parameterization of these models for more extreme bio-geographical regions such as subarctic climates has not been performed and relies on assuming that less extreme Boreal parameterization will be applicable. In this study, conditions at Svanhovd for the growing seasons 2018/19 shall serve as a reference for present and probable future conditions in northern Fennoscandia.
It is important to understand the implications of using these more generic parameterizations for more extreme climate regions. Here we explore how risk estimates might be over or underestimated using generic vs bespoke subarctic parameterization. We also explore the role of \chem{[O_3]} variability and changing climate variables to understand how this risk might alter in the future with the advent of climate change. This allows an understanding of how the threat from ozone may change over time and which are the key aspects of plant physiology that might determine risk. Hence, we determine which parameters require special attention to define accurately for pollution impact assessment and more generally for describing gas exchange response to changing climates that will influence biochemical cycling. This might be one key to solve long-standing issues of coupled earth system models (ESM) which often lack PFTs specifically representing subarctic conditions in their land surface models \citep{GMD:Poulter2015,JAMES:Lawrence2019}.\\

@@

There are two global reanalysis products available at the European Centre for Medium-Range Weather Forecasts (ECMWF) that include atmospheric tracers, in particular ozone, the Monitoring Atmospheric Composition and Climate (MACC) and the newer Copernicus Atmosphere Monitoring Service reanalysis (CAMSRA) \citep{ACP:Inness2013, ACP:Inness2019}. The temporal as well as spatial resolution of these reanalysis products is rather coarse: 3-hourly and $0.75\,\unit{^\circ}\times0.75\,\unit{^\circ}$ or roughly $29.3\,\unit{km}\times 83.4\unit{km}$ at the location of Svanvik. From Copernicus member institutions, higher resolution surface ozone reanalysis data are available for Europe (CAMS Regional Air Quality (CAMSRAQ) reanalysis ensemble). The CAMSRAQ is based on nine European state-of-the-art numerical air quality models. The ensemble mean is at fairly high spatial and temporal resolution compared to the global reanalysis: $0.1\,\unit{^\circ}\times 0.1\,\unit{^\circ}$ or roughly $3.9\,\unit{km}\times 11.1\,\unit{km}$ and 1-hourly, respectively. The periods covered differ and no data is available before the turn of the millennium. The CAMSRA and CAMSRAQ are available in near real time, but only the former covers a period of sufficient length for climate analysis (2003 -- present). All reanalyses use different but, at that time, the latest versions of the operational weather forecast system (OpenIFS) from ECMWF. They differ substantially in the assimilated observational ozone data. The MACC reanalysis assimilates only satellite-derived tropospheric column ozone, while CAMSRA also includes ozone profiles from satellite retrievals. Most importantly for our study, in situ observations from ozone near-surface station networks are only assimilated in the CAMSRAQ reanalysis ensemble. The quality concerning observations in northern Fennoscandia will be discussed in Section~\ref{subsubsec:clim_ozone}. Due to the found substantial discrepancies between reanalysis data and site observations in northern Fennoscandia, we refrain from using any of the global reanalysis data.

\begin{table}[t]
  \caption{Global/regional ozone reanalysis products used in this study.}
  \label{tab:ozone_rea}
\begin{tabular}{llcclcc}
\tophline
Name & Provider & \multicolumn{2}{c}{Resolution} & Time period & Meteorological forcing & \chem{O_3} assimilation\\
&              & spatial & temporal & & \\
\middlehline
MACC & ECMWF & $0.75^\circ \times 0.75^\circ$ & 3-hourly & 2003 -- 2012 & OPS & satellite $^\triangledown$\\
CAMSRA & ECMWF & $0.75^\circ \times 0.75^\circ$ & 3-hourly & 2003 -- 2012 $^\dagger$ & ERA5 / OPS $^\ddagger$ & satellite $^\blacktriangledown$\\
CAMSRAQ & Copernicus & $0.1^\circ \times 0.1^\circ$ & 1-hourly & 2014 -- 2018 $^\dagger$ & OPS $^\star$ & in situ $^\vartriangle$\\
\bottomhline
\end{tabular}
\belowtable{$^\dagger$ Subset of the data used in this study. The full data set extends into near real-time; $^\ddagger$ ERA5 (2003--2016), OPS (later); $^\star$ EURAD uses WRF for downscaling of operational IFS; $^\triangledown$ MLS, OMI - tropospheric column; $^\blacktriangledown$ SCIAMACHY, MIPAS, MLS, OMI, GOME2, SBUV2 - tropospheric column + profile; $^\vartriangle$ METEO France NRT.} % Table Footnotes
\end{table}

@@

We hypothesize that perennial vegetation will likely prioritize their growth or fast adolescence and hence their carbon sequestration, during a short, subarctic growing season. Because the mapping manual version of the $\mathrm{DO_3SE}$ model currently does not simulate $A_\mathrm{n}$, we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$. We further assume that $f_\mathrm{VPD}$ and $f_\mathrm{SWP}$ suffice our biomes. The procedure is graphically illustrated in Fig.~\ref{fig:f_temp_grassland} for grassland. We aim to increase the enclosed area between probability density function (PDF) and the respective response function. We show two climatological reference periods 1990s and 2000s to indicate that the vegetation is subject to ongoing climate change and hence continuous acclimation. Resulting response functions for Norway spruce (Picea abies) and downy birch are displayed in Appendix~Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}.

@@

As pointed out by \citet{ACP:Bueker2012}, a very important factor in deriving a proper $\mathrm{POD_y}$ is the soil texture or how much water the soil can hold. As Svanhovd is located in the bed of the Pasvik river, we assume a sandy loam texture at $60\,\unit{cm}$ depth. The estimate of leaf boundary layer resistance is a function of wind speed which will be influenced by the tree height and leaf dimensions. Leaves of downy birch in northern Fennoscandia are smaller than southern specimens' with leaf dimensions of $(3.0\pm 0.5)\,\unit{cm}$. In addition, \citet[][p.~52]{NINA2004} indicate an average tree height of $13.5\,\unit{m}$ in the area. Both these factors will affect leaf boundary layer resistance, the smaller leaves would tend to reduce leaf boundary layer resistances whilst the shorter trees would reduce canopy roughness and increase canopy boundary layer resistances.

@@

From the difference between 2018 and 2019, we infer a substantial interannual variability. The interannual variability is estimated to be larger than our reported systematic deviations and can be of the same order of magnitude as the effect of the temperature acclimation. This means that the choice of temperature acclimation has the potential to either revoke or amplify the risk of ozone damage for vegetation. However, the risk of ozone damage in downy birch and perennial grassland in comparison with the generic default MM parameterizations is increased, regardless.

@@

\begin{acknowledgements}
  Bj{\o}rg Rognerud for processing SeNorge.no data with respect for begin of growing season.
  Tore Flatlandsmo Berglen for hourly pressure data from Svanhovd.
  This work was financed by the Research Council of Norway (NRC) grand No. 268073.
\end{acknowledgements}
